---
title: 'Spatial Analysis: NASA Meteorite-Landings'
author: "Patrick Reza Schnurbusch"
date: "08/01/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(readr)
library(rmarkdown)
library(leaflet)
library(leaflet.providers)
library(leaflet.extras)

library(spatialEco)
library(sp)

library(highcharter) 

# Set highcharter options
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))

meteorites <- read.csv("./meteorite-landings.csv", encoding = "UTF-8")
```

```{r}
summary(meteorites)
```

The dataset contains the following variables:

* name: the name of the meteorite (typically a location, often modified with a number, year, composition, etc)
id: a unique identifier for the meteorite

* nametype: one of:
  * valid: a typical meteorite
  * relict: a meteorite that has been highly degraded by weather on Earth
  
* recclass: the class of the meteorite; one of a large number of classes based on physical, chemical, and other characteristics (see the Wikipedia article on [meteorite classification](https://en.wikipedia.org/wiki/Meteorite_classification) for a primer)

* mass: the mass of the meteorite, in grams

* fall: whether the meteorite was seen falling, or was discovered after its impact; one of:
  * Fell: the meteorite's fall was observed
  * Found: the meteorite's fall was not observed

* year: the year the meteorite fell, or the year it was found (depending on the value of fell)
* reclat: the latitude of the meteorite's landing
* reclong: the longitude of the meteorite's landing
* GeoLocation: a parentheses-enclose, comma-separated tuple that combines reclat and reclong

# Data Cleaning 

We'll clean out incorrect reccomended lat/long coordinates, we'll also filter for USA via using a bounding box for lat/long within the most north/south/east/west bounds of the continental US. 


```{r}
meteorites <- meteorites %>% 
  dplyr::filter(year >= 860 & year <= 2016 ) %>%   # filter out weird years
  dplyr::filter(reclat != 0 | reclong != 0) %>% 
  dplyr::arrange(desc(year)) %>% 
  dplyr::filter(!is.na(mass)) %>% 
  dplyr::filter(mass > 0) %>% 
  dplyr::mutate(year = as.numeric(year))

# src for bbox 
# https://gist.github.com/graydon/11198540

meteorites_clean <- meteorites %>% 
  dplyr::filter(reclat <= 49.3457868 & reclat >= 24.7433195) %>% #filter for north and south 
  dplyr::filter(reclong >=  -124.7844079 & reclong <= -66.9513812) # filter for east and west 
```


## Meteorites by Mass

```{r, fig.width = 8}
pal <- colorNumeric(palette = "RdYlBu", domain = meteorites_clean$mass, reverse = T)

leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircleMarkers(lng = ~reclong, lat = ~reclat,
                   color = ~pal(mass), 
                   radius = 3, 
                   stroke = FALSE, fillOpacity = 0.5) %>% 
  
  addLegend("bottomright", pal = pal, values = ~mass,
            title = "Mass (grams)",
            labFormat = labelFormat(suffix = " g"),
            opacity = 1
  )
```


## Heatmap of Meteorite impacts 

```{r}
leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addHeatmap(lng = ~reclong, lat = ~reclat, 
             gradient = "viridis", 
             minOpacity = 0.1, 
             cellSize = 1, 
             radius = 3, blur = 15)
```


## Meteorites by Observed Status 

Here we'll visualize the meteorites via whether they were spotted on the ground after impact ("found") or while in the mid-fall ("fell)

Overwhelmingly we can see most meteorites were spotted while on the ground after-impact was made, we have modified the size argument to help make the quantity more acute on our map. 
```{r, fig.width = 8}

pal <- colorFactor(c("#6EC5E9", "#FF1C51"), domain = c("Fell", "Found"))

leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircleMarkers(lng = ~reclong, lat = ~reclat,
                   color = ~pal(fall), 
                   radius = ~ifelse(fall == "Found", 1, 3),
                   stroke = FALSE, fillOpacity = 1) %>% 
  
  addLegend("bottomright", pal = pal, values = ~fall,
            title = "Found Status",
            opacity = 1)
```

Frequency distribtion by year 

```{r}
hchart(meteorites_clean$year,
          color = "Purple", 
          name = "year") %>% 
  
  hc_yAxis(title = list(text = "Frequency (n)")) %>% 
  hc_xAxis(title = list(text = "<b> Year Observed </b>")) %>% 
  hc_add_theme(hc_theme_monokai())
```

## Meteorites via Classification  

Examining wether 


# Hot Spot Spatial Analysis 

Useful good [here](https://www.publichealth.columbia.edu/research/population-health-methods/hot-spot-spatial-analysis)

## Test for spatial autocorrelation/clustering in data

```{r}
xy <- meteorites_clean[,c(9,8)]

meteorites_clean <- SpatialPointsDataFrame(coords = xy, data = meteorites_clean,
                               proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

print(nni(meteorites_clean))
```


Our NNI score is less than 1, we can consider this as a good indication of clustering and can reject the Null Hypothesis that the points are uniform in pattern

## Create the hotspot map
## Define the hotspot map legend threshold
