---
title: 'Point Pattern Analysis: NASA Meteorite-Landings'
author: "Patrick Reza Schnurbusch"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
      theme: darkly
      code_folding: hide
---

```{r setup, include=FALSE}
library(dplyr)
library(readr)
library(rmarkdown)
library(leaflet)
library(leaflet.providers)
library(leaflet.extras)

library(spatialEco)
library(sp)
library(sf)

library(spatstat)

library(highcharter) 

# for downloading US state shp files 
library(tigris)


library("USAboundaries")

# Set highcharter options
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))

meteorites <- read.csv("./meteorite-landings.csv", encoding = "UTF-8")
```



This dataset was obtained from Kaggle [here](https://www.kaggle.com/nasa/meteorite-landings) and includes the latitude and longitude of meteorite impacts. This dataset includes 45,000 meteorite imapacts, in this analysis we will be examining those impacting the United States specifically. 

Our goal is to apply point-pattern analysis and areal statistics to this dataset. 

The dataset also contains the following variables:

* name: the name of the meteorite (typically a location, often modified with a number, year, composition, etc)
id: a unique identifier for the meteorite

* nametype: one of:
  * valid: a typical meteorite
  * relict: a meteorite that has been highly degraded by weather on Earth
  
* recclass: the class of the meteorite; one of a large number of [meteorite classifications](https://en.wikipedia.org/wiki/Meteorite_classification) based on physical, chemical, and other characteristics 

* mass: the mass of the meteorite (grams)

* fall: 
  * Fell - the meteorite's was observed before impact midfall
  * Found: the meteorite's was observed after impact

* year: the year the meteorite fell, or the year it was found (depends on the value of fell)
* reclat: the latitude of the meteorite's landing
* reclong: the longitude of the meteorite's landing
* GeoLocation: a parentheses-enclose, comma-separated value of the reclat & reclong

```{r}
# data cleaning 

meteorites <- meteorites %>% 
  dplyr::filter(year >= 860 & year <= 2016 ) %>%   # filter out weird years
  dplyr::filter(reclat != 0 | reclong != 0) %>% 
  dplyr::arrange(desc(year)) %>% 
  dplyr::filter(!is.na(mass)) %>% 
  dplyr::filter(mass > 0) %>% 
  dplyr::mutate(year = as.numeric(year))

meteorites_clean <- meteorites %>% 
  dplyr::filter(reclat <= 49.3457868 & reclat >= 24.7433195) %>% #filter for north and south 
  dplyr::filter(reclong >=  -124.7844079 & reclong <= -66.9513812) # filter for east and west 
```

# Exploratory Analysis  

## Meteorites visualized by mass

```{r, fig.width = 8}
pal <- colorNumeric(palette = "RdYlBu", domain = meteorites_clean$mass, reverse = T)

leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircleMarkers(lng = ~reclong, lat = ~reclat,
                   color = ~pal(mass), 
                   radius = 3, 
                   stroke = FALSE, fillOpacity = 0.5) %>% 
  
  addLegend("bottomright", pal = pal, values = ~mass,
            title = "Mass (grams) <hr>",
            labFormat = labelFormat(suffix = " g"),
            opacity = 1
  )
```


## Heatmap of Meteorite impacts 

```{r}
leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addHeatmap(lng = ~reclong, lat = ~reclat, 
             gradient = "Spectral", 
             minOpacity = 0.1, 
             cellSize = 1, 
             radius = 3, blur = 15)
```


## Meteorites by Observed Status 

Here we'll visualize the meteorites via whether they were spotted on the ground after impact ("found") or while in the mid-fall ("fell)

Overwhelmingly we can see most meteorites were spotted while on the ground after-impact was made, we have modified the size argument to help make the quantity more acute on our map. 
```{r, fig.width = 8}

pal <- colorFactor(c("#6EC5E9", "#FF1C51"), domain = c("Fell", "Found"))

leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircleMarkers(lng = ~reclong, lat = ~reclat,
                   color = ~pal(fall), 
                   radius = ~ifelse(fall == "Found", 1, 3),
                   stroke = FALSE, fillOpacity = 1) %>% 
  
  addLegend("bottomright", pal = pal, values = ~fall,
            title = "Found Status",
            opacity = 1)
```


Distribution of Meteorites by Years 

```{r}
# Create a continuous palette function
pal <- colorBin(
  palette = "Spectral",bins = 5,
  domain = meteorites_clean$year, reverse = T)


leaflet(data = meteorites_clean, height=500, width=910) %>% 
  addFullscreenControl() %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>% 
  addCircleMarkers(lng = ~reclong, lat = ~reclat,
                   color = ~pal (year), 
                   radius = 3,
                   stroke = FALSE, fillOpacity = 1) %>% 
  
  addLegend("bottomleft", 
            colors =c("#2B83BA", "#ABDDA4", "#FFFFBF","#FDAE61","#D7191C"),
            labels = c("1600 - 1700", "1700 - 1800", "1800 - 1900", "1900 - 2000", "2000 - 2016"), 
            title = "Year Found",
            opacity = 1)
```

Most of the meteorites were discovered within last century (1900 - 2000)

Frequency distribtion by year 

```{r}
hchart(meteorites_clean$year,
          color = "Purple", 
          name = "year") %>% 
  
  hc_yAxis(title = list(text = "Frequency (n)")) %>% 
  hc_xAxis(title = list(text = "<b> Year Observed </b>")) %>% 
  hc_add_theme(hc_theme_monokai())
```



## Density based analysis

### Quadrat Density 

We'll divide the observation into a grid (h = 5, width = 10) and count the points that lie within each quadrat 

```{r warning=FALSE}
# get the window for the data, i.e. the boundaries 
W <- owin(xrange = c(min(meteorites_clean$reclong), 
                     max(meteorites_clean$reclong)),
          
          yrange = c(min(meteorites_clean$reclat), 
                     max(meteorites_clean$reclat)))

# select just the lat/long columns 
pp1 <- meteorites_clean %>% 
  dplyr::select(reclong, 
                reclat)

# convert to ppp object 
meteorites_ppp <- as.ppp(pp1, W = W)

Q <- quadratcount(meteorites_ppp, nx= 5, ny=10)

# plot results 
plot(meteorites_ppp, pch=20, cols="#FF1C51", main=NULL)  # Plot points, with the tallies 
plot(Q, add=TRUE)  # Add quadrat grid
```


### Computing the density 
```{r}

meteorites_ppp.km <- rescale(meteorites_ppp, 1000, "km")

# Compute the density for each quadrat (in counts per km2)
Q   <- quadratcount(meteorites_ppp.km, nx= 6, ny=3)
Q.d <- intensity(Q)

# Plot the density

plot(intensity(Q, image=TRUE), main=NULL, las=1)  # Plot density raster
plot(meteorites_ppp.km, pch=20, cex=0.6, col="#39FF14", add=TRUE)  # Add points
```

### Kernel density raster

```{r}
K1 <- density(meteorites_ppp.km) # Using the default bandwidth
plot(K1, main=NULL, las=1)
contour(K1, add=TRUE)
```


## Distance based analysis

We'll apply three difference measures for distance bases analysis: 

* The average nearest neighbor 
* K & L function 
* pair correlation function $g$

### Average nearest neighbor analysis (ANN)

Computing the ANN distances between imapct sites. Compute the average *first* nearest neighbor distance (km), set `k = 1`

```{r}
# set k - 1 for the first nearest neighbor distance 
mean(nndist(meteorites_ppp.km, k = 1))
```

Compute the average *second* nearest neighbor, set `k = 2`
```{r}
mean(nndist(meteorites_ppp.km, k = 2))

```
### ANN Plot 
```{r}
ANN <- apply(nndist(meteorites_ppp.km, k=1:2000),2,FUN=mean)
# generate an ANN vs neighbor order plot
plot(ANN ~ eval(1:2000), type="b", main="ANN Plot", las=1)
```

The bottom axis shows the neighbor order number and the left axis shows the average distance (km). We can see a linear trend via

### K & L Functions 

Computing the K function

```{r warning=FALSE}
K <- Kest(meteorites_ppp.km)
plot(K, main=NULL, las=1)
```

This plot is visualizing the different estiamtes of $K$ depending on the edge correction. The edge corrections plotted are listed in the legend. 

$K_{iso}$ refers to an *isotropic* implementation

$K_{trans}$ refers to an *translate* implementation

$K_{bord}$ refers to an *border * implementation

$K_{pois}$ refers to an *border * implementation


### Kernel density adjusted for covariate

Here we'll estimate the meteorite point process intensity using the mass 

In the following example, a Starbucks store point process’ intensity is estimated following the population density raster covariate







## Hypothesis testing 

Running an average nearest neighbor analysis for impact sites, in this we assume a *uniform point density across space* 

```{r}
ann.p <- mean(nndist(meteorites_ppp.km, k =1))
ann.p
```

The observed average nearest neighbor distance is 0.0002431465 km (0.24 m).

Generate the distribution of an expected 







# Hot Spot Spatial Analysis 

Useful good [here](https://www.publichealth.columbia.edu/research/population-health-methods/hot-spot-spatial-analysis)

## Test for spatial autocorrelation/clustering in data

```{r}
xy <- meteorites_clean[,c(9,8)]

meteorites_clean <- SpatialPointsDataFrame(coords = xy, data = meteorites_clean,
                               proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

print(nni(meteorites_clean))
```


Our NNI score is less than 1, we can consider this as a good indication of clustering and can reject the Null Hypothesis that the points are uniform in pattern

## Create the hotspot map
## Define the hotspot map legend threshold







# Areal Statistics 

## Meteorite Landings by State 

In this section we'll tally the number of meteorites via the State in which they land to gather some areal statistics. 


```{r,echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
# Downloading US State polygons 
states <- states(cb = TRUE)

# Convert to sf data
points <- st_as_sf(meteorites_clean)
# Set CRS 
states <- st_transform(x = states, crs = 4326)
points <- st_transform(x = points, crs = 4326)


# Intersection between polygon and points ---------------------------------
intersection <- st_intersection(x = states, y = points)
# Count number ofintersections per polygon (i.e. states)
states$pt_count <- lengths(st_intersects(states, points))
```


```{r, message=FALSE}
# Interactive Map 
pal <- colorNumeric(
  palette = "viridis",domain = states$pt_count, reverse = F, alpha = T)

labels_info = sprintf(
  "<h2> %s </h2>
  <hr>
  Meteorites: %s ",
  states$NAME, 
  states$pt_count
) %>% lapply(htmltools::HTML)

leaflet(states) %>%
  addProviderTiles("CartoDB.DarkMatter") %>%
  addPolygons(fillColor = ~pal(pt_count), 
              weight = 2,
              opacity = 0.25,
              color = "white",
              dashArray = "1",
              fillOpacity = 0.6,
              
              highlight = highlightOptions(
                weight = 3,
                color = "#8ffcff",
                fillColor = "#8ffcff",
                dashArray = "1",
                fillOpacity = 0.6,
                bringToFront = TRUE), 
              
              label = labels_info, 
              labelOptions = labelOptions(
                style = list("font-family" = "arial",
                             "font-style" = "bold",
                             "color" = "#FF1C51",
                             "background" = "black",
                             "border-color" = "#FFD300")
                )
              ) %>%
  
  setView(-98.5795, 39.8282, zoom=3) %>% 
  
  addLegend("bottomright", 
            pal = pal, 
            values = ~pt_count,
            title = "Meteorites <hr>",
            opacity = 1)
```


```{r}
sfno <- st_as_sf(mapsimple)
st_crs(sfno)
carto_ref <- cartogram(states, "pt_count")
```





