---
title: "Metorite Landings"
author: "Patrick Reza Schnurbusch"
date: "05/04/2021"
output: 
  html_document:
    theme: spacelab
    highlight: default
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Amelia)
library(dplyr)
library(plotly)

library(leaflet)
library(leaflet.providers)
library(leaflet.extras)

library(RColorBrewer)
library(sf)

meteorites <- read.csv("./meteorite-landings.csv", encoding = "UTF-8")
```

# Preamble 

This dataset was obtained from Kaggle [here](https://www.kaggle.com/nasa/meteorite-landings) and includes the lat/lon coordinates of meteorite impact sites.  
There 45,000 impact sites and in this analysis we will be analyzing the ones inside of the continential United States. 

# Data Cleaning 

First we need to check for data integrity. If we count the NA's we can see that they are most frequent in the lat/long, year, and then mass.

```{r}
colSums(is.na(meteorites))
```

We can confirm this visually by passing our dataset to a missing values map from the 'Amelia' package. 

```{r}
missmap(meteorites, 
        main = "Missingness Map of Meteorite Impact Dataset",
        legend = T, 
        y.labels = NULL, 
        y.at = NULL,
        col = c("#ff6961", "#84d9ff"))
```

If we calculate the percentage of missing values in relation to all values (i.e. product = # of rows * # of columns) we get approx 3.2% missing. 

```{r}
sum(is.na(meteorites))/prod(dim(meteorites))
```

This number is low enough where we can simply filter these data-points out of our dataset and proceed without the need over imputation. We remove the missing values from our dataset and run a missing values map to confirm this as shown below. 

```{r}
# Filtering out missing values 
meteorites_clean <- meteorites %>% 
  dplyr::filter(!is.na(year) & !is.na(mass) & !is.na(reclat) & !is.na(reclong)) 

missmap(meteorites_clean, 
        main = "Missingness Map of Meteorite Impact Dataset",
        legend = T, 
        y.labels = NULL, 
        y.at = NULL,
        col = c("#ff6961", "#84d9ff"))

```

If we run `summary()` We can see below two (2) main issues: 

```{r}
summary(meteorites_clean)
```

First, We can see some meteorites have a mass of 0 grams, this could be because the actual mass was so small it did not register past a decimal point. To correct this we can filter these objects out. 

```{r}
meteorites_clean <- meteorites_clean %>% 
  dplyr::filter(mass != 0)
```


2. We can also see that our data starts at 601 and ends at 2101. We can assume these to be errors, and will need to filter out. 

We will use 2016 as our 'ceiling' or the max point we want to find out. We can remove these rows and then make a density plot to see where our 'floor' should, that is to mean the lowest values we want filtered out. 

```{r}
meteorites_clean <- meteorites_clean %>% 
  dplyr::filter(year <= 2016)
 
density <- density(meteorites_clean$year)
plot_ly(x = ~density$x, y = ~density$y, type = 'scatter', mode = 'lines', fill = 'tozeroy') %>% 
  layout(xaxis = list(title = 'Year'),
         yaxis = list(title = 'Density'))
```
We can see the number of meteorite impacts increases over time as both our understanding and technology improves, however we see a significant jump approx after 1960's. With this in mind we'll filter out for values before 1960, this choice is arbitrary admittedly but will give us a clearer picture of 'recent' meteorite impact sites in our research area. 

```{r}
meteorites_clean <- meteorites_clean %>% 
  dplyr::filter(year >= 1960)
```

Next we need to filter out for this impacting the continental united states. To achieve this we'll need to select the points for the north/south and east/west boundaries of the continental 


```{r}

# hacky method using the northern/southern, and eastern/western most coords to create a box to filter on 
meteorites_usa <- meteorites_clean %>% 
  dplyr::filter(reclat <= 49.3457868 & reclat >= 24.7433195) %>% #filter for north and south 
  dplyr::filter(reclong >=  -124.7844079 & reclong <= -66.9513812) # filter for east and west 
```

With our contemporary US state boundary file we will need to convert our dataframe to an sf object to perform the operation. 

Now with out dataset properly cleaned and prepped we can move onto the Point-Pattern-Analysis (PPP) step. 

# Point Pattern Analysis

First we'll map out all 971 points on a map to see how the data appears. 

## Mass

### Impacts by Mass

Here we have mapped out the impact sites, color coding and scaling the size of each point by mass (g).  

```{r, fig.width = 8}
pal <- colorNumeric(palette = "RdBu", domain = meteorites_usa$mass, reverse = T)

leaflet(data = meteorites_usa, height=500, width=910) %>% 
        addFullscreenControl(position = "topright") %>% 
        addResetMapButton() %>% 
        addProviderTiles(providers$Stamen.TonerLite)%>% 
        addCircleMarkers(lng = ~reclong, lat = ~reclat,
                         color = ~pal(mass), 
                         radius = meteorites_usa$mass/1e5, 
                         stroke = FALSE, fillOpacity = 1) %>% 
        
        addLegend("bottomright", pal = pal, values = ~mass,
                  title = "Mass (grams) <hr>",
                  labFormat = labelFormat(suffix = " g"),
                  opacity = 1)
```
We can see some clustering along the border between New Mexico & Texas but mostly it's sporadic and random.   

## Impact Mass Distribution 

Below we have boxplotted the points to gauge the frequency. We can see that the mass clusters below 50K grams, with the median being ~300g. 
```{r}
plot_ly(meteorites_usa) %>% 
        add_boxplot(y = ~mass, jitter = 0.5, pointpos = -1.8, boxpoints = 'all', 
                    marker = list(color = '#36C5F0'), line = list(color = '#FFFC00'), name = "mass(g)") %>% 
        layout(yaxis = list(title = 'mass(g)')) %>% 
        
        layout(plot_bgcolor='black') %>% 
        layout(paper_bgcolor='black') 
```

## Heatmap by Mass of Impact
Here we have produced a heatmap to more clearly illustrate clusters for impact sites. We can see clearly that our original site is indeed a cluster but also that along California/Arizona is also one for consideration. 

```{r}
leaflet(data = meteorites_usa, height=500, width=910) %>% 
        addFullscreenControl(position = "topright") %>% 
        addResetMapButton() %>% 
        addProviderTiles(providers$Stamen.TonerLite) %>% 
        addHeatmap(lng = ~reclong, lat = ~reclat, 
                   gradient = "magma", 
                   minOpacity = 0.1, 
                   cellSize = 1, 
                   radius = 3, blur = 15)

```

### Meteorites by Observed Status 

In our dataset meteorites are coded via whether they were spotted on the ground after impact ("found") or while in the mid-fall ("fell)

Overwhelmingly we can see most meteorites were spotted while on the ground after-impact was made, we have modified the size argument to help make the quantity more acute on our map. 
```{r, fig.width = 8}
pal <- colorFactor(c("#FF1C51", "#006AFF"), domain = c("Fell", "Found"))

labels <- c("Mid-Air", "Ground")

leaflet(data = meteorites_usa, height=500, width=910) %>% 
        addFullscreenControl(position = "topright") %>% 
        addResetMapButton() %>% 
        addProviderTiles(providers$Stamen.TonerLite) %>% 
        addCircleMarkers(lng = ~reclong, lat = ~reclat,
                         color = ~pal(fall), 
                         radius = ~ifelse(fall == "Found", 2, 4),
                         stroke = FALSE, fillOpacity = 1) %>% 
        
        addLegend("bottomright", pal = pal, values = ~fall,
                  title = "<u>Found Status</u>",
                  opacity = 1, 
                  labFormat = function(type, cuts, p){
                          paste0(labels)})
```

## Distribution of Meteorites by Years 

Here we can interestingly that the more recent impact sites are now impacting California/Arizona and that the datapoints along New Mexico /Arizona are mostly 'older' impact sites. 
```{r}
pal <- colorBin(
        palette = "RdBu",bins = 5,
        domain = meteorites_usa$year, reverse = T)


leaflet(data = meteorites_usa, height=500, width=910) %>% 
        addFullscreenControl(position = "topright") %>% 
        addResetMapButton() %>% 
        addProviderTiles(providers$Stamen.TonerLite) %>% 
        addCircleMarkers(lng = ~reclong, lat = ~reclat,
                         color = ~pal (year), 
                         radius = 3,
                         stroke = FALSE, fillOpacity = 1) %>% 
        
        addLegend("bottomright", pal = pal, values = ~year,
                  title = "<u>Year Found</u>",
                  labFormat = labelFormat(),
                  opacity = 0.9)
```

## Impacts over Time 

```{r}
plot_ly(data = meteorites_usa,
        x = ~year,
        type = "histogram",
        cumulative = list(enabled=TRUE), 
        marker = list(color = '#006AFF'), 
        name = "Cumulative") %>%
        
        add_trace(data = meteorites_usa,
                  x = ~year,
                  type = "histogram", 
                  cumulative = list(enabled=F), 
                  marker = list(color = '#00ffff'),
                  name = "Per Year") %>% 
        
        layout(title = "Meteorite Impacts in America (1960 - 2016)",
               titlefont = list(
                       family = "Agency FB", size = 45, color = '#ffffff'),
               font = list(family = "Agency FB", size = 25),
               margin = 10,
               paper_bgcolor='black',
               plot_bgcolor='black',
               xaxis = list(title = "Year", 
                            color = '#ffffff'),
               yaxis = list(title = "No. of Impacts", 
                            color = '#ffffff'))

```

# Meteorite Classifications

one of a large number of [meteorite classifications](https://en.wikipedia.org/wiki/Meteorite_classification) based on physical, chemical, and other characteristics 

```{r}
# Create a frequency for our class variable 
class_freq_table <- meteorites_usa %>% 
        dplyr::group_by(recclass) %>% 
        dplyr::tally() %>% 
        dplyr::ungroup() %>% 
        dplyr::arrange(desc(n)) %>% 
        dplyr::top_n(n = 10, wt = n)


colors =  brewer.pal(10, "PuOr")

plot_ly(class_freq_table, labels = ~recclass, values = ~n, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        
        marker = list(colors = colors,line = list(color = 'white', width = 2)),
        showlegend = FALSE, 
        hoverinfo = 'text',
        text = ~paste('<b>Meteorite Classification:</b>', recclass, "<br>", 
                      "<b>No. of Impacts:</b>", n, "</br>", 
                      "<b>No. of Impacts:</b>", round(n/sum(n)*100, 1),"%", "</br>")) %>% 
        
        layout(title = 'Meteorites Classes in USA (1960 - 2016)',
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)) %>% 
        
        layout(titlefont = list(
                       family = "Agency FB", size = 45, color = '#ffffff'),
               font = list(family = "Agency FB", size = 25),
               margin = 10,
               paper_bgcolor='black',
               plot_bgcolor='black')
```

